---
title: "p8105_hw2_ra2965"
output: github_document
---

Read and clean the Mr. Trash Wheel sheet:

```{r Problem1 trash wheel sheet}
library(tidyverse)

trash_wheel = readxl::read_excel ("/Users/anran/Desktop/DataScience/week4/hw2/data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", sheet = "Mr. Trash Wheel") %>%
  
janitor::clean_names() %>%

select(-x15) %>%
  
drop_na(dumpster) %>%

mutate(sports_balls = as.integer(sports_balls))
```
Read and clean precipitation data for 2017 and 2018:

```{r Problem1 precipitation sheet}
precipitation_2018 = readxl::read_excel("/Users/anran/Desktop/DataScience/week4/hw2/data/HealthyHarborWaterWheelTotals2018-7-28.xlsx",sheet = "2018 Precipitation", skip = 1) %>%
janitor::clean_names() %>%
drop_na(.) %>%
mutate(year = 2018)


precipitation_2017 = readxl::read_excel("/Users/anran/Desktop/DataScience/week4/hw2/data/HealthyHarborWaterWheelTotals2018-7-28.xlsx",sheet = "2017 Precipitation",skip = 1) %>%
janitor::clean_names() %>%
drop_na(.) %>%
mutate(year = 2017)

precipitation = union_all(precipitation_2017, precipitation_2018) %>%
mutate(month = month.name[month])

```
There are `r nrow(trash_wheel)` observations in the Mr. Trash Wheel dataset and `r nrow(precipitation)` observations in the precipitation dataset. For available data, the total precipitation in 2018 is `r sum(pull(precipitation_2018, total))`. The median number of sports balls in a dumpster in 2017 is `r median(pull(trash_wheel, sports_balls))`.

```{r Problem2 pols-month}
pols_month = read_csv(file = "../data/fivethirtyeight_datasets/pols-month.csv") %>%
janitor::clean_names() %>%
separate("mon", into = c("year","month","day"), sep = "-") %>%
mutate(month = month.name[as.numeric(month)]) %>%
  select(-prez_dem,-prez_gop)


```

