---
title: "p8105_hw2_ra2965"
output: github_document
---
#Problem 1

##Read and clean the Mr. Trash Wheel sheet:

```{r Problem1 trash wheel, message = FALSE}
library(tidyverse)

trash_wheel = readxl::read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = "Mr. Trash Wheel") %>%
  janitor::clean_names() %>%
    select(-x15, -x16,-x17) %>%
    drop_na(dumpster) %>%
    mutate(sports_balls = round(sports_balls,0)) %>%
  mutate(sports_balls = as.integer(sports_balls))
```

## Read and clean precipitation data for 2017 and 2018, and then combine them:

```{r  Problem1 precipitation, message = FALSE}

precipitation_2018 = readxl::read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",sheet = "2018 Precipitation", skip = 1) %>%
janitor::clean_names() %>%
drop_na(.) %>%
mutate(year = 2018)


precipitation_2017 = readxl::read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",sheet = "2017 Precipitation",skip = 1) %>%
janitor::clean_names() %>%
drop_na(.) %>%
mutate(year = 2017)

precipitation = full_join(precipitation_2017, precipitation_2018) %>%
mutate(month = month.name[month])

```
There are `r nrow(trash_wheel)` observations and `r ncol(trash_wheel)` variables in the Mr. Trash Wheel dataset. The dataset includes key variables of `r colnames(trash_wheel)`, which are the dumpter number, date of collection, litter type and amount of litter for each type.


There are `r nrow(precipitation)` observations and `r ncol(precipitation)` variables in the precipitation dataset. The dataset includes key variables of `r colnames(precipitation)`, which are the precipitation data for each month in 2017 and 2018.

For available data, the total precipitation in 2018 is `r sum(pull(precipitation_2018, total))`. The median number of sports balls in a dumpster in 2017 is `r median(pull(trash_wheel%>% filter(year == 2017), sports_balls))`.

#Problem 2

##Read and clean the data file pols-month:
```{r Problem2 pols-month, message = FALSE}

pols = read_csv(file = "./data/fivethirtyeight_datasets/pols-month.csv") %>%
janitor::clean_names() %>%
separate(mon, into = c("year","month","day"), sep = "-", convert = TRUE) %>%
mutate(month = month.abb[month]) %>%
  mutate(president = ifelse(prez_gop > 0, "gop","dem")) %>%
  select(-prez_gop, -prez_dem, -day)

```

##Read and clean the data file snp:

```{r Problem2 snp,message = FALSE}

snp = read_csv(file = "./data/fivethirtyeight_datasets/snp.csv") %>%
separate("date", into = c("month","day","year"), sep = "/", convert = TRUE) %>%
  arrange(year, month)%>%
  mutate(month = month.abb[month])%>%
  select(year, month, everything(),-day)

```


##Read and clean the data file unemployment:
```{r Problem2 unemployment, message = FALSE}

unemployment = read_csv(file = "./data/fivethirtyeight_datasets/unemployment.csv") %>% pivot_longer(
             Jan:Dec,
             names_to = "month",
             values_to = "percent_unemployment"
) %>% select(year = Year, everything())

```

##Join the datasets：

```{r joined datasets, message = FALSE}

snp_pols = left_join(pols, snp, by = c("month","year"))
snp_pols_emp = left_join(snp_pols, unemployment, by = c("month","year"))

```

The cleaned file “pols-month” contains `r nrow(pols)` observations of `r ncol(pols)` variables. The variables are related to the number of national politicians who are democratic or republican at any given time during `r min(pull(pols, year))` to `r max(pull(pols, year))`. 

Key variables:

president: an indicator of whether the president was republican on the associated date. 

gov_gop, sen_gop, rep_gop: the number of republican governors, senators and representatives on the associated date. 

gov_dem,sen_dem, rep_dem: the number of democratic governors, senators and representatives on the associated date.


The cleaned file “snp” contains `r nrow(snp)` observations of `r ncol(snp)` variables. The year range in this dataset is from `r min(pull(snp, year))` to `r max(pull(snp, year))`.

Key variable: 

close: the closing values of the S&P stock index on the associated date, which is often used as a representative measure of stock market as a whole.


The file “unemployment” contains `r nrow(unemployment)` observations of `r ncol(unemployment)` variables. The year range in this dataset is from `r min(pull(unemployment, year))` to `r max(pull(unemployment, year))`. 

Key variable:

percent_unemployment: the percentage of unemployment of the associated month and year.


The joined file snp_pols_emp which chombines the three datadets contains `r nrow(snp_pols_emp)` observations of `r ncol(snp_pols_emp)` variables. The year range in this dataset is from `r min(pull(snp_pols_emp, year))` to `r max(pull(snp_pols_emp, year))`. 

Key variables:

gov_gop, sen_gop, rep_gop: the number of republican governors, senators and representatives on the associated date.

gov_dem,sen_dem, rep_dem: the numbers for democratic overnors, senators and representatives on the associated date. 

president: an indicator of whether the president was republican on the associated date.

close: the closing values of the S&P stock index on the associated date. 

percent_unemployment: the percentage of unemployment of the associated date.

#Problem 3

##Import and clean the datafile of popular baby names:

```{r Problem3 babynames, message = FALSE}
baby_names = read_csv(file = "./data/Popular_Baby_Names.csv") %>%
  janitor::clean_names() %>%
  mutate(ethnicity = recode(ethnicity, "ASIAN AND PACI" = "ASIAN AND PACIFIC ISLANDER", "BLACK NON HISP"= "BLACK NON HISPANIC","WHITE NON HISP" = "WHITE NON HISPANIC"))%>%
  mutate(childs_first_name = str_to_lower(childs_first_name)) %>%
  distinct()

```

##Produce two reader-friendly tables:

```{r female and male name, message = FALSE}
baby_names %>% filter(childs_first_name == "olivia",gender == "FEMALE") %>% select(childs_first_name, year_of_birth,ethnicity, rank) %>%
  pivot_wider(names_from = "year_of_birth",
              values_from = "rank") %>% knitr::kable()


baby_names %>% filter(gender == "MALE",rank == 1) %>% select(childs_first_name, year_of_birth,ethnicity) %>% pivot_wider(names_from = "year_of_birth",
              values_from = "childs_first_name") %>% knitr::kable()
```

##Produce a scatter plot:
```{r scatterplot, message = FALSE}
library(ggridges)
plot_df = filter(baby_names, ethnicity == "WHITE NON HISPANIC", gender == "MALE", year_of_birth == 2016)
ggplot(plot_df, aes(x = rank, y = count)) + geom_point() +
  labs(
    title = "The first name of male, white non-hispanic children born in 2016",
    x = "The rank in popularity of that name (rank)",
    y = "The number of children with a name (count)")
  
```

