---
title: "p8105_hw2_ra2965"
output: github_document
---

Read and clean the Mr. Trash Wheel sheet:

```{r Problem1 trash wheel}
library(tidyverse)

trash_wheel = readxl::read_excel("/Users/anran/Desktop/DataScience/week4/hw2/data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = "Mr. Trash Wheel") %>%
  
janitor::clean_names() %>%

select(-x15, -x16,-x17) %>%
  
drop_na(dumpster) %>%

mutate(sports_balls = as.integer(sports_balls))
```
Read and clean precipitation data for 2017 and 2018:

```{r Problem1 precipitation}
precipitation_2018 = readxl::read_excel("/Users/anran/Desktop/DataScience/week4/hw2/data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",sheet = "2018 Precipitation", skip = 1) %>%
janitor::clean_names() %>%
drop_na(.) %>%
mutate(year = 2018)


precipitation_2017 = readxl::read_excel("/Users/anran/Desktop/DataScience/week4/hw2/data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",sheet = "2017 Precipitation",skip = 1) %>%
janitor::clean_names() %>%
drop_na(.) %>%
mutate(year = 2017)

precipitation = union_all(precipitation_2017, precipitation_2018) %>%
mutate(month = month.name[month])

```
There are `r nrow(trash_wheel)` observations in the Mr. Trash Wheel dataset and `r nrow(precipitation)` observations in the precipitation dataset. For available data, the total precipitation in 2018 is `r sum(pull(precipitation_2018, total))`. The median number of sports balls in a dumpster in 2017 is `r median(pull(trash_wheel, sports_balls))`.


Second, clean the data in snp.csv using a similar process to the above. For consistency across datasets, arrange according to year and month, and organize so that year and month are the leading columns.

Third, tidy the unemployment data so that it can be merged with the previous datasets. This process will involve switching from “wide” to “long” format; ensuring that key variables have the same name; and ensuring that key variables take the same values.

Join the datasets by merging snp into pols, and merging unemployment into the result.

Write a short paragraph about these datasets. Explain briefly what each dataset contained, and describe the resulting dataset (e.g. give the dimension, range of years, and names of key variables).

Note: we could have used a date variable as a key instead of creating year and month keys; doing so would help with some kinds of plotting, and be a more accurate representation of the data. Date formats are tricky, though. For more information check out the lubridate package in the tidyverse.




```{r Problem2 pols-month}

pols_month = read_csv(file = "../data/fivethirtyeight_datasets/pols-month.csv") %>%
janitor::clean_names() %>%
separate(mon, into = c("year","month","day"), sep = "-", convert = TRUE) %>%
mutate(month = month.name[month]) %>%
  mutate(president = ifelse(prez_gop > 0, "gop","dem")) %>%
  select(-prez_gop, -prez_dem, -day)

```

```{r Problem2 snp}

snp = read_csv(file = "../data/fivethirtyeight_datasets/snp.csv") %>%
separate("date", into = c("day","month","year"), sep = "/") %>%
  arrange(year, month)%>%
  mutate(month = month.name[as.numeric(month)])%>%
  select(year, month, everything(),-day)

```

```{r Problem2 unemployment}
unemployment = read_csv(file = "../data/fivethirtyeight_datasets/unemployment.csv") %>%
pivot_longer(Jan:Dec,
             names_to = "month",
             values_to = "percent_unemployment"
) 


```

